---
output:
  html_document: default
  pdf_document: default
---

## SEURAT WRAPPER TO REPORT SIGNIFICANT MARKERS AND REACTOME PATHWAYS

A new Seurat user is generally interested to run the single cell pipeline with less effort, quickly look at the expressed markers in different clusters and identify the pathways of interest within each cluster. This repository is a simple R package that has 4 functions:

[1]. SeuratPreprocess Function to convert Counts Data to Normalized and Scaled Seurat object for highly variable genes.

[2]. SeuratLowDim Function to convert Scaled Seurat Object from [1] to object that has clusters identified and data transformed to visualize in 2d (ie, PCA followed by t-SNE and UMAP). 

[3]. SeuratMarkers Function to convert seurat object to all markers list, along with significant markers (based on minimum percent of cells in the cluster).

[4]. ReactomeData Function to convert seurat object to get the pathways identified using ReactomeGSA R package.

```{r cars}
library(Seurat)
library(ReactomeGSA)
library(scReport)
```

```{r}
# Loading sample counts data (Mouse cell atlas)
mca.matrix <- readRDS(file = "data/MCA_merged_mat.rds")
mca.metadata <- read.csv(file = "data/MCA_All-batch-removed-assignments.csv", row.names = 1)
mca.matrix.10K <- mca.matrix[,1:1000]
```

## [1]. Function to convert Counts Data to Normalized and Scaled Seurat object for highly variable genes.
## [2]. Function to convert Scaled Seurat Object from [1] to object that has clusters identified and data transformed to visualize in 2d (ie, PCA followed by t-SNE and UMAP).
```{r SeuratPreprocess and SeuratLowDim functions}
scaled_seurat_object <- SeuratPreprocess(mca.matrix.10K)

low_dim_object <- SeuratLowDim(scaled_seurat_object)
```

## [3]. The Seurat markers list is slightly modified to pick a threshold for pct.1 and pct.2 i.e., the minimum number of cells where the gene of interest is seen.
## [4]. Next, we use Reactome GSA R package to identify the pathways of interest using the cluster information of Seurat object.

```{r}

Reactome_pathways_object <- ReactomeData(low_dim_object)
```

## Here is a quick glimpse of the output objects and tables outputted.
```{r}
scaled_seurat_object
```

```{r}
low_dim_object
```

```{r}
Reactome_pathways_object[1]
```



```{r}
Reactome_pathways_object[3]
```

Pathways of interest:

```{r}
plot_gsva_pathway(Reactome_pathways_object[[1]], pathway_id = rownames(Reactome_pathways_object[[3]])[1])
```

Same pathways showed as a heatmap:
```{r}
plot_gsva_heatmap(Reactome_pathways_object[[1]], max_pathways = 15)
```

Relevant pathways based on user's list:
```{r}

relevant_pathways <- c("R-HSA-983170", "R-HSA-388841", "R-HSA-2132295", "R-HSA-983705", "R-HSA-5690714")
plot_gsva_heatmap(Reactome_pathways_object[[1]], 
                  pathway_ids = relevant_pathways,
                  dendrogram = "col", 
                  scale = "row", # scale for each pathway
                  key = FALSE, 
                  lwid=c(0.1,4)) 
```

